\myTitle

### 2024-03-04 body-text-of-paper.tex
These are edits that Rachael sent me
Text and fig caps in Results to end of mykiss stuff.
Just have the pigs left to do in the results.

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, with number of chromosomes and sex-averaged recombination rates
for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
admixture fraction of each simulated individual as the cumulative length of all genome segments
that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

These permutations are implemented with the genotypes provided by the user.
\gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
$2L$ columns (two for each locus). If the data have been phased, then the phased alleles from the
first haplotype should occupy the
first entry for each locus within an individual, while alleles phased together on the second haplotype should occupy the
second entry for each locus.
There can be several populations represented in the data set, but
each individual must belong to exactly one
population.

@@

As expected, the perceived ability to estimate $q_A$ with ADMIXTURE was higher
for larger values of $L$ and for smaller values of $N$ (Figure~\ref{fig:bias-sims}).
%%%
\begin{figure*}
\newcommand{\biassimscap}{\footnotesize ADMIXTURE estimates of $q_A$ as described in
{\em An introductory motivating example}. In these results, the apparent ability to estimate $q_A$
is a bias resulting from the use of sampling with replacement to simulate new, admixed genotypes to
test in ADMIXTURE.  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
"ref-2" and "ref-1" show the $Q$-values estimated by ADMIXTURE for the reference samples
(even though they are treated by the program as being of known origin). ({\bf b}) Results for the
unsupervised analysis in ADMIXTURE. In both {\bf a} and {\bf b}, the different columns represent different numbers of markers
from 10 to 100,000, while different rows represent different original sample sizes, $N$, taken. Colors
of the boxplots indicate how many new individuals, $n$, of each $q_A$ value were simulated during each
replicate.}
\includegraphics[width=0.96\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
\caption[\biassimscap]{\biassimscap}
\label{fig:bias-sims}
\end{figure*}
%%%
It is interesting to note
that, for numbers of loci,  $L$, that are 100 (or smaller) the effect is observed only when
sample size, $N$, is as small as 25---and even then the effect is slight.  However, when the
number of markers increases to 100,000---values that are becoming commonplace with the availability of chip-
or sequencing-based approaches---it is apparent that the sampling-with-replacement approach
induces an extreme bias.  With 100,000 markers, even if sample sizes ($N$) are as large as 250 individuals,
using na\"{i}ve simulations improperly indicates that admixture fractions of individuals can be
accurately estimated, even when there are no genetic differences, whatsoever between the populations that
samples $A$ and $B$ are drawn.

@@

When hybrids were simulated in a way that accounts for the physical linkage of
markers, the variance in admixture fractions was considerably higher than when
markers were simulated under the assumption that none of them are physically
linked. (Fig.~\ref{fig:ad-fract-sims}).
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
This is, however, not the case.  The bottom panel of Fig.\ref{fig:ad-fract-sims} shows the distribution
of the true admixture fraction (calculated as the total length of genome from each of the two
different species) of each individual when simulated with physical linkage using \gscramble{}.
When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
$BC_4$ all have admixture fraction distributions that overlap greatly.  This shows that,
in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
between different backcross categories on the basis of their admixture fractions.

@@

As expected, there was no RISPI apparent in the ADMIXTURE-estimated
ancestry proportions  when simulating individuals
using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
were around 0.5, which is expected when the samples are drawn from the same
population.
%%%
\begin{figure}
\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
fractions, $Q$, for individuals simulated from samples from the same population,
simulated using \gscramble{}.  There is a considerable variation in the estimates,
but the central tendency of them all is around 0.5, as would be expected in this case
within which there is not genetic basis for distinguishing individuals from the different
samples.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
in which the RISPI is readily apparent.}
\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
\caption[\gssimcap]{\gssimcap}
\label{fig:norispi}
\end{figure}
%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories.

%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT, $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
{\sc NewHybrids} does can yield a poor estimate of the certainty around different inferences.

>>>>

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, with 18 chromosomes (2.35 gigabases of physical genome length)  and
sex-averaged recombination rates
for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
admixture fraction of each simulated individual as the cumulative length of all genome segments
that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

These permutations are implemented with the genotypes provided by the user.
\gscramble{} accepts genetic data as a matrix with $N$ rows (one for each individual) and
$2L$ columns (two alleles for each locus). If the data have been phased, then the phased alleles from the
first haplotype should occupy the
first entry for each locus within an individual, while alleles phased together on the second haplotype should occupy the
second entry for each locus.
There can be several populations represented in the data set, but
each individual must belong to exactly one
population.

@@

As expected, the perceived ability to estimate $q_A$ with ADMIXTURE was higher
for larger values of $L$ and for smaller values of $N$ (Figure~\ref{fig:bias-sims}).
%%%
\begin{figure*}
\newcommand{\biassimscap}{\footnotesize ADMIXTURE estimates of $q_A$ as described in
{\em An introductory motivating example}. In these results, the apparent ability to estimate $q_A$
is a bias resulting from the use of sampling with replacement to simulate new, admixed genotypes to
test in ADMIXTURE (e.g., these simulated samples suffer from RISPI).  ({\bf a}) Results from the supervised analysis in ADMIXTURE, where 
"ref-2" and "ref-1" show the $Q$-values estimated by ADMIXTURE for the reference samples
(even though they are treated by the program as being of known origin). ({\bf b}) Results for the
unsupervised analysis in ADMIXTURE. In both {\bf a} and {\bf b}, the different columns represent different numbers of markers
from 10 to 100,000, while different rows represent different original sample sizes, $N$, taken. Colors
of the boxplots indicate how many new individuals, $n$, of each $q_A$ value were simulated during each
replicate.}
\includegraphics[width=0.96\textwidth]{figures/bias-sims-unsup-and-sup.pdf}
\caption[\biassimscap]{\biassimscap}
\label{fig:bias-sims}
\end{figure*}
%%%
It is interesting to note
that, for numbers of loci,  $L$, that are 100 (or smaller) the effect is observed only when
sample size, $N$, is as small as 25---and even then the effect is slight.  However, when the
number of markers increases to 100,000---values that are becoming commonplace with the availability of chip-
or sequencing-based approaches---it is apparent that the sampling-with-replacement approach
induces an extreme bias.  With 100,000 markers, even if sample sizes ($N$) are as large as 250 individuals,
using na\"{i}ve simulations improperly indicates that admixture fractions of individuals can be
accurately estimated, even when there are no genetic differences, whatsoever between the populations that
samples $A$ and $B$ are drawn.

@@

When hybrids were simulated in a way that accounts for the physical linkage of
markers, the variance in admixture fractions was considerably higher than when
markers were simulated under the assumption that none of them are physically
linked (Fig.~\ref{fig:ad-fract-sims}).
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{} assuming a genetic map with 18 chromosomes spanning 2.35 gigabases with recombination fractions like those found in pigs
\citep{tortereau2012high}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
This is, however, not the case.  The bottom panel of Fig.~\ref{fig:ad-fract-sims} shows the distribution
of the true admixture fraction (calculated as the total length of genome from each of the two
different species) of each individual when simulated with physical linkage using \gscramble{}.
When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
$BC_4$ all have admixture-fraction distributions that overlap greatly.  This shows that,
in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
between different backcross categories on the basis of their admixture fractions.

@@

As expected, there was no RISPI apparent in the ADMIXTURE-estimated
ancestry proportions  when simulating individuals
using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
were around 0.5, which is expected when the samples are drawn from the same
population.
%%%
\begin{figure}
\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
fractions, $Q$, for individuals simulated from samples from the same population,
simulated using \gscramble{}.  There is considerable variation in the estimates,
but the central tendency of them all is around 0.5, as would be expected in this case
within which the different
samples are drawn from the same population.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
in which the RISPI is readily apparent.}
\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
\caption[\gssimcap]{\gssimcap}
\label{fig:norispi}
\end{figure}
%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT, $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
{\sc NewHybrids} does can yield a poor estimate of the uncertainty around different inferences.
